---
title: "Session 3 Visualization"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives and Learning

- Difference between Exploratory vs. Explanatory Analysis
- Vector Graphic (Java Script) vs. PNG (ggplot) type of graphics


## Tools and Packages

- ggplot2
- dygraphs and family of htmlwidgets
- plotly 


## Using ggplot2 for exploratory analysis

- ggplot (grammer of graphics) 
- (1) Works well with dplyr and tidyr
- (2) Large Selection of Graphics 
- (3) Sharp and Layered Graphics 
- (4) Big negative is that graphic is not interactive (there are ways around this!)

## Visulaizing statistical inference

- Basic ggplot graph
```{r, echo=TRUE,message=FALSE}
load("/home/data/session1.RData")
load("/home/data/session2.RData")
# install.packages("ggplot2")
require(ggplot2)
require(dplyr)
prices %>% ggplot(aes(x=as.Date(date,format="%Y-%m-%d"),y=price.usd)) + geom_point(stat="identity",position="identity") 


```

- Little more visual analysis
```{r, echo=TRUE, message=FALSE}
require(dplyr)
prices %>% ggplot(aes(x=as.Date(date,format="%Y-%m-%d"),y=price.usd)) + geom_point(aes(group=factor(year),colour=factor(year))) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("price") + xlab("date")+ geom_smooth(aes(group=factor(year)),method = "lm")
 
```

- Different type of Plots

```{r, echo=TRUE, message=FALSE}
require(dplyr)
prices %>% ggplot(aes(x=year,y=price.cad)) + geom_boxplot(aes(group=year),outlier.colour = "red") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("price") + xlab("date")
 
```

- Finding the relationship between CAD/USD Prices to BTC

```{r, echo=TRUE, message=FALSE}
require(dplyr)
prices %>% ggplot(aes(x=price.cad,y=price.usd)) + geom_point(aes(group=factor(year),colour=factor(year)))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("USD") + xlab("CAD") + geom_smooth(aes(group=factor(year)),method = "lm")
 
```

- Extending ggplot2 
[Source]("http://www.r-bloggers.com/ggplot2-cheatsheet-for-visualizing-distributions/")

```{r, echo=TRUE, message=FALSE}
require("gridExtra")
empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     )



scatter <- prices %>% ggplot(aes(price.cad, price.usd)) + 
  geom_point(aes(color=factor(year))) + 
  scale_color_manual(values = c("orange", "purple","green")) + 
  theme(legend.position=c(1,1),legend.justification=c(1,1)) 

#marginal density of x - plot on top
plot_top <- prices %>% ggplot(aes(price.cad, fill=factor(year))) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c("orange", "purple","green")) + 
  theme(legend.position = "none")

#marginal density of y - plot on the right
plot_right <- prices %>% ggplot(aes(price.usd, fill=factor(year))) + 
  geom_density(alpha=.5) + 
  coord_flip() + 
  scale_fill_manual(values = c("orange", "purple","green")) + 
  theme(legend.position = "none") 

#arrange the plots together, with appropriate height and width for each row and column
grid.arrange(plot_top, empty, scatter, plot_right, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4))

```

## Interactive Graphics using Java Script based plots

- Collection of Java Script Based Graphics including D3
- Good Source from [HTML Widgets]("http://www.htmlwidgets.org/showcase_leaflet.html")
- Simple Time Series Analysis Using Dygraphs
- More Information on the package: [Link]("http://rstudio.github.io/dygraphs/")

```{r, echo=FALSE,message=TRUE}
library(dygraphs)
dygraph(prices.xts[,c("date","price.cad","price.usd")], main = "USD/BTC and CAD/BTC") %>%
  dySeries("price.cad", label = "CAD/BTC") %>%
  dySeries("price.usd", label = "USD/BTC") %>%
  dyRangeSelector(dateWindow = c("2015-01-01", "2016-05-01"))

```

- Layering Analytics using TTR Package 

```{r, echo=FALSE,message=TRUE}
library(dygraphs)
library(TTR)
tmp.value<-prices.xts[,c("price.usd","price.usd.ret")]
tmp.value<-na.omit(tmp.value)
tmp.value$max<-runSD(tmp.value$price.usd.ret,n=20)*1.6+SMA(tmp.value$price.usd,n=50)
tmp.value$min<--runSD(tmp.value$price.usd.ret,n=20)*1.6+SMA(tmp.value$price.usd,n=50)

dygraph(tmp.value[,c("price.usd","max","min")], main = "USD/BTC Price and Vol") %>%
  dySeries(c("min","price.usd","max"), label = "USD/BTC") %>%
  dyOptions(colors = RColorBrewer::brewer.pal(3, "Set1"))

```

- ggplot interactive plots using plotly package

```{r, echo=FALSE,message=TRUE}
library(plotly)
p<-prices %>% ggplot(aes(x=price.cad,y=price.usd)) + geom_point(aes(group=factor(year),colour=factor(year)))+ theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylab("USD") + xlab("CAD") + geom_smooth(aes(group=factor(year)),method = "lm")
ggplotly(p)

```



