---
title: "Session 2 Data Munging and Cleaning"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives and Learning

- Clean Data and Make it Ready for Analysis
- tidy format and semi-raw data
- time series data 

## Tools and Packages

- dplyr
- tidyr  
- xts format


## Using dplyr to clean data  

- Loading environment

```{r, echo=TRUE,message=FALSE}
load("/home/data/session1.RData")
```

- Group_by and Summarise 

```{r,echo=TRUE,message=FALSE}
require(dplyr)
tmp<-prices %>% group_by(year) %>% summarise(count=n())
head(tmp)
tmp<-prices %>% group_by(year) %>% summarise_each(funs(n(),mean),matches("price"))
head(tmp)
tmp<-prices %>% group_by(year) %>% summarise(max=max(price.usd.ret))
head(tmp)
tmp<-prices %>% na.omit() %>% group_by(year) %>% summarise(max=max(price.usd.ret))
head(tmp)
tmp<-prices %>% na.omit() %>% group_by(year) %>% summarise(sd=sd(price.usd.ret))
head(tmp)
```

- filter data

- Filter with numeric data
```{r, echo=TRUE,message=FALSE}
tmp<-prices %>% filter(abs(price.usd.ret) > 100)
head(tmp)
```

- Filter by date and arranging dates
```{r, echo=TRUE,message=FALSE}
tmp<-prices %>% filter(date > as.Date("2015-01-01",format="%Y-%m-%d")) %>% dplyr::arrange(-desc(date))
head(tmp)
tail(tmp)
```

## Pivot Table Functionality in R and tidyr

```{r, echo=TRUE,message=FALSE}
#install.packages(tidyr)
require(tidyr)
tmp<-prices %>% dplyr::select(date,year,price.usd.ret)%>%
           tidyr::spread(year,price.usd.ret)
head(tmp)
tail(tmp)
tmp<-tmp %>% gather(year,price,-date)
head(tmp)
str(tmp)
```

## Apply Functions using Purrr Package

- Easy Example
```{r, echo=TRUE,message=FALSE}
# install.packages("purrr")
require(purrr)
tmp<-tmp %>% map_if(is.factor,as.character)
str(tmp)
```
- More complicated problem
```{r, echo=TRUE,message=FALSE}
prices %>%
  split(.$year) %>%
  map(~ lm(price.cad.ret ~ price.usd.ret, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")

prices %>%
  split(.$year) %>%
  map(~ lm(price.cad ~ price.usd, data = .)) %>%
  map(summary) %>%
  map_dbl("r.squared")
```

Further reading: [purrr package github](https://github.com/hadley/purrr)

## Using xts vs. dataframe Pros/Cons

- xts/ts - Pros: Easier Date Functionality & Access to Ecnometric Packages
cons: does not work well with data.frame based packages dplyr and ggplot
- data.frame - Pros: Fast & works well with dplyr / Cons: more difficult for date munipilation
- use both depending functionality and package
- conversion between df and xts

```{r, echo=TRUE,message=FALSE}
require(xts)
head(prices)
str(prices)
prices.xts<-xts(prices[,-1],order.by=as.Date(prices[,1],format="%Y-%m-%d"))
str(prices.xts)
head(prices.xts)
ret<-diff(prices.xts[,c("price.cad","price.usd")])
ret<-na.omit(ret)
summary(lm(price.usd~price.cad,data=ret))
save(prices.xts,file="/home/data/session2.RData")
```

## Further Learning and Links

- Wrangle Chapter - Hadley Wickham [R for Data Science](http://r4ds.had.co.nz/)
- Courses 1-2 - Johns Hopkins Data Science Specialization [Coursera](https://www.coursera.org/specializations/jhu-data-science)

## Other Useful Packages:

- quantmod: Time series financial data analysis  
- TTR: Technical Trading and Quant Signals
- quantstrat: Full functional blotter and Quant Trading Systems
- PerformanceAnalytics: Risk Management and Financial Performance Package





