---
title: "Session 5 Statistical Learning"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives and Learning

- Tools Descriptive Statistics
- Inference and Hypothesis Testing

## Tools and Packages

- stats package
- time series
- twitteroutlierdetection
- caret package


## Simple descriptive statistics

```{r, echo=FALSE, message=FALSE}
load("/home/data/session2.RData")
summary(prices.xts)
head(zoo::rollmean(prices.xts,10))
```

## Regresssion and Statistical Inference

```{r, echo=FALSE, message=FALSE}
load("/home/data/session2.RData")
summary(prices.xts)
head(zoo::rollmean(prices.xts,10))
```

## Time Series Forcasting

- Univariant Time Series

```{r, echo=FALSE, message=FALSE}
#install.packages("forecast")
require(forecast)
load("/home/data/session1.RData")
fit <- arima(prices$price.usd, order=c(1, 1, 0))
fit <- auto.arima(prices$price.usd)
library(forecast)
forecast(fit, 5)
plot(forecast(fit, 30))
```

- Seasonal Forecat


```{r, echo=FALSE, message=FALSE}
fit <- HoltWinters(prices$price.usd, beta=FALSE, gamma=FALSE)
# double exponential - models level and trend
fit <- HoltWinters(prices$price.usd, gamma=FALSE)
# triple exponential - models level, trend, and seasonal components
#fit <- HoltWinters(prices$price.usd)
# predict next three future values
library(forecast)
out<-data.frame(forecast(fit, 30),date=index(forecast(fit, 30)))
plot(forecast(fit, 30))

```


## Outlier Detection 

```{r, echo=TRUE,message=FALSE}
#install.packages("devtools")
#devtools::install_github("twitter/AnomalyDetection")
#install.packages("devtools")
#devtools::install_github("twitter/BreakoutDetection")
library(BreakoutDetection)
library(AnomalyDetection)
value<-prices[,c("date","price.usd")]
value$date<-as.POSIXlt(value$date,format="%Y-%m-%d")
res = AnomalyDetectionTs(value,max_anoms=0.1,direction='both',plot=TRUE)
res$plot

names(value)<-c("timestamp","count")
res = breakout(value, min.size=24, method='multi', beta=.001, degree=1, plot=TRUE)
res$plot

```

## Machine Learning

```{r, echo=TRUE, message=FALSE}

l.1<-lag(prices.xts$price.usd.ret,n=1);names(l.1)<-"L1";prices.xts<-merge(prices.xts,l.1)
l.2<-lag(prices.xts$price.usd.ret,n=2);names(l.2)<-"L2";prices.xts<-merge(prices.xts,l.2)
l.3<-lag(prices.xts$price.usd.ret,n=3);names(l.3)<-"L3";prices.xts<-merge(prices.xts,l.3)
l.5<-lag(prices.xts$price.usd.ret,n=5);names(l.5)<-"L5";prices.xts<-merge(prices.xts,l.5)
l.10<-lag(prices.xts$price.usd.ret,n=10);names(l.10)<-"L10";prices.xts<-merge(prices.xts,l.10)
tmp.stg<-data.frame(prices.xts,row.names=NULL)
```

```{r, echo=TRUE, message==FALSE, eval=FALSE}
require(caret)
tmp.stg<-na.omit(tmp.stg)
featurePlot(x=tmp.stg[,-7],y=tmp.stg$price.usd.ret,type=c("g","p","smooth"))
forTraining<-createDataPartition(y=tmp.stg$price.usd.ret,p=3/4)[[1]]
trainingSet<-tmp.stg[forTraining,]
testSet<-tmp.stg[-forTraining,]
set.seed(975)

modFormula<-paste("price.usd.ret ~ L1+L2+L3+L5+L10")
modFormula<-as.formula(modFormula)
controlObject<-trainControl(
                      method="repeatedcv",
                      repeats=5,
                      number=10)

lmModel<-train(modFormula,data=trainingSet,method="lm",trControl=controlObject)

svmModel<-train(modFormula,data=trainingSet,
                 method="svmRadial",
                 preProc=c("center","scale"),
                 trControl=controlObject)


nnetGrid<-expand.grid(.decay=c(0.001,0.01,0.1),.size=seq(1,27,by=2),.bag=FALSE)
#nnetModel<-train(modFormula,data=trainingSet,
#                 method="avNNet",
#                 tuneGrid=nnetGrid,
#                 preProc=c("center","scale"),
#                 trControl=controlObject,
#                 linout=TRUE,
#                 trace=FALSE,
#                 maxit=1000
#                 )

rpartModel<-train(modFormula,data=trainingSet,
                 method="rpart",
                 tuneLength=30,
                 trControl=controlObject)

allResamples<-resamples(list("Linear"=lmModel,"SVM"=svmModel,"CART"=rpartModel))

caret::parallelplot.resamples(allResamples,metric = "Rsquared")

```

## Further Learning and Links

- Max Kuhn [Applied Predictive Modelling](http://appliedpredictivemodeling.com/)
- [Caret Package](http://topepo.github.io/caret/index.html)
- Courses on Statistical Inference and Practical Machine Learning - Johns Hopkins Data Science Specialization [Coursera](https://www.coursera.org/specializations/jhu-data-science)





