---
title: "Intro to Data Science for Risk Managers and Trading Professionals"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Why Data Science?

- Risk Managers and Trading Professionals = Decision Engineers
- Data Science is a science behind getting to better decisions
- Data Science is about generating hypothesis and testing
- Data Science is about identifying causal relationship between different market factors

## Our Objective

- Risk Managers can benefit greatly from the world of Data Science
- Data Science from a the Risk Manager/Trading Professional perspective
- The objective is to give a small glimpse into possibilities 
- This is going to introductory and 101 and we hope to build on this through community feedback

## What is some of the use cases in Risk Management

- Outlier Detections and Regime Switching
- Generating Stress Scenarios
- Visualizing Market Relationships
- Developing Prediciton Models

## Data Science Approach 

- Data Science is about uncovering unknowns
- Experimentation and Validated Learing
- Lean Approach and (1) Build (2) Measure (3) Learn Cycle
- Fast cycle development requires fit-for-purpose tools

## Tools and Resources

- Cloud Platform
- Scalable and flexible data environment
- Raw Data at your fingertips is 80% of the work
- Open Source tools with large user community

## What do I use and Why?

- R is one of the most popular tools of Data Scientists
- R is open source with large user community
- Single interface and many tools
- Rich statistical library sets

## How to use R and Setup for the sessions

- Power of R is in Packages
- RStudio IDE: [www.rstudio.com]

## Data Science Approach and Risk Management Example

- Ingesting Data from an API 
- Cleaning Data 
- Exploratory Visualization
- Descriptive Statistics 
- Predictive Models 
- Deploy the tools as an online app

## BitCoin/Crypto Currency Example

- BitCoins potentially new asset class 
- BitCoins are completely new to me no previous experience and bias
- We have three currency pairs (1) CAD/BTC (2) US/BTC and (3) EUR/BTC how do we use Data Science to learn about this market and possibly develop a trading model
- This problem is more about journy and learning than outcome


## Flavour of what is to come!

- Ingesting Data From APIs and Stageing for Use
```{r cars, echo = TRUE}
summary(cars)
```

## Data Viz as Exploratory tool

```{r pressure}
plot(pressure)
```

