---
title: "Session 5 Web apps"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("/home/data/session1.RData")
load("/home/data/session2.RData")
# install.packages("ggplot2")
require(ggplot2)
require(dplyr)
library(dygraphs)
require("gridExtra")
```

## Objectives and Learning

- Reproducible Analysis R Markdown
- Shiny Package and Interactive Apps
- Deploying Web Apps in the Cloud

## Tools and Packages

- Rstudio IDE 
- knitr
- Shiny

## Concept of RMarkdown and Interactive Documents

- Convergence of Web and Document
- App first strategy - it is marginally more work
- Writing in single source and rendering multiple formats - Simple Example from RStudio

## Shiny and Web Apps 

- Powerful Web Application without knowing HTML and Java Script
- Concept of UI and Server Files
- Bootstrap UI interface and scaffolding
- Concept of Reactivity 
- UI Interactive vs. Server Interactive 

## Developing a Simple App

This R Markdown document is made interactive using Shiny. This is an easy alternative to develop and deploy shiny web interactive analytics 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

```{r}
input.choices<-c("price.cad","vol.cad","price.usd","vol.usd","price.usd.ret","price.cad.ret") 

inputPanel(
  selectInput("value", label = "Value 1",
              choices = input.choices, selected =c("date","price.cad","price.usd"),multiple = T)
)

renderDygraph({
dygraph(prices.xts[,c(input$value)], main = "example") %>%
#  dySeries("price.cad", label = "CAD/BTC") %>%
#  dySeries("price.usd", label = "USD/BTC") %>%
  dyRangeSelector(dateWindow = c("2015-01-01", "2016-05-01"))
})

  
```




```{r eruptions, echo=F}

input.choices<-c("price.cad","vol.cad","price.usd","vol.usd","price.usd.ret","price.cad.ret") 
  
  
inputPanel(
  selectInput("value1", label = "Value 1",
              choices = input.choices, selected = "price.cad"),
  selectInput("value2", label = "Value 2",
              choices = input.choices, selected = "price.usd")
)

renderPlot({
  
empty <- ggplot()+geom_point(aes(1,1), colour="white") +
     theme(                              
       plot.background = element_blank(), 
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.border = element_blank(), 
       panel.background = element_blank(),
       axis.title.x = element_blank(),
       axis.title.y = element_blank(),
       axis.text.x = element_blank(),
       axis.text.y = element_blank(),
       axis.ticks = element_blank()
     )

prices.tmp<-prices[,c(input$value1,input$value2,"year")]
names(prices.tmp)<-c("value.1","value.2","year")

scatter <- prices.tmp %>% ggplot(aes(value.1, value.2)) + 
  geom_point(aes(color=factor(year))) + 
  scale_color_manual(values = c("orange", "purple","green")) + 
  theme(legend.position=c(1,1),legend.justification=c(1,1)) 

#marginal density of x - plot on top
plot_top <- prices.tmp %>% ggplot(aes(value.1, fill=factor(year))) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c("orange", "purple","green")) + 
  theme(legend.position = "none")

#marginal density of y - plot on the right
plot_right <- prices.tmp %>% ggplot(aes(value.2, fill=factor(year))) + 
  geom_density(alpha=.5) + 
  coord_flip() + 
  scale_fill_manual(values = c("orange", "purple","green")) + 
  theme(legend.position = "none") 

#arrange the plots together, with appropriate height and width for each row and column
grid.arrange(plot_top, empty, scatter, plot_right, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4))
  
})
```

## Deploying Shiny Apps in the Cloud

- shinyapp.io
- local sharing
- shiny server/shiny server pro


## Further Learning and Links

- [RStudio/Shiny](http://shiny.rstudio.com/)
- [Bootstrap](http://getbootstrap.com/)




