---
title: "Session 2 Data Ingesting and Stageing"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Objectives and Learning

- Source Raw Data from Web
- Stage Data for Easy Use
- Saving Data for Future use

## Tools and Packages

- dplyr
- rjson
- httr
- quandl 
- mongolite

## Installing R Packages

- Power of R in Packages
- Over 7000 Packages in CRAN
- CRAN/Github best Sources for R packages

Easy one Step!

```{r eval=FALSE,echo=TRUE,message=FALSE}
install.packages("xts")
```

## Different Data Types in R

- numeric
```{r,echo=TRUE,message=FALSE}
x<-5
str(x)
```
- character
```{r,echo=TRUE,message=FALSE}
x<-"g"
str(x)
x<-c("a","b","c")
str(x)
```

- data.frame
```{r,echo=TRUE,message=FALSE}
x.df<-data.frame(a=c("v1","v2"),b=c(1,2),stringsAsFactors = F)
str(x.df)
```
- list
```{r,echo=TRUE,message=FALSE}
x<-list(v3=x.df,v4=c(1,2,3))
str(x)
```


## Using Web APIs 

- JSON Format
```{r,echo=TRUE,message=FALSE}
#install.packages("jsonlite")
require(jsonlite)
samplefile<-fromJSON("https://www.quandl.com/api/v3/datasets/GDAX/CAD.json")
str(samplefile)
value<-samplefile$data$data
```
- Using Quandl Package

```{r,echo=TRUE,message=FALSE}
#install.packages("Quandl")
require(Quandl)
cad<-Quandl("LOCALBTC/CAD")
usd<-Quandl("LOCALBTC/USD")
str(cad)
head(cad)
```

## Very Brief Introduction to dplyr Package

- Best tool for Data Munging for data.frame 
- documentation and references: [Introduction to dplyr](https://github.com/hadley/dplyr) 
- dplyr cheat sheet [Link](chrome-extension://oemmndcbldboiebfnladdacbdfmadadm/https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)

- example: Merge USD/BTC and CAD/BTC data by date sets calcuate day over day change for price and volume

```{r, echo=TRUE,message=FALSE}
# install.packages("dplyr")
#install.packages("lubridate")
require(dplyr)
require(lubridate)
cad.clean<-cad %>% dplyr::select(date=Date,price.cad=Last,vol.cad=matches("Volume")) 

usd.clean<-usd %>% dplyr::select(date=Date,price.usd=Last,vol.usd=matches("Volume"))
prices<-cad.clean %>% inner_join(usd.clean, by="date")
prices<-prices %>% mutate(price.usd.ret=price.usd-lag(price.usd)) %>%
           mutate(price.cad.ret=price.cad-lag(price.cad)) %>% 
           mutate(year=year(date),month=month(date))
head(prices)

```

## Databases and Persistance

- Many choice among SQL and NoSQL Application
- Database as a datasource
- Persistance of Analytics Data
R Packages: DBI, RMySQL, rmongodb, or mongolite

```{r, echo=TRUE, message=TRUE}
# install.packages("mongolite")
require(mongolite)
m<-mongo(collection="price",db="garp",url="mongodb://localhost")
m$drop()
m<-mongo(collection="price",db="garp",url="mongodb://localhost")
m$insert(prices)
prices<-m$find()

```

More resources: [Link](https://cran.r-project.org/web/packages/mongolite/vignettes/intro.html)

## Saving Your Work

- Writing to CSV
- Saving R Session

```{r, echo=TRUE,message=FALSE}
# install.packages("readr")
require(readr)
prices %>% readr::write_csv("/home/data/prices.csv") 
head(readr::read_csv("/home/data/prices.csv"))
save(prices,file="/home/data/session1.RData")
rm(list=ls(all=TRUE)) 
load("/home/data/session1.RData")
head(prices)
```






